---
title: "607 project 1"
author: "Jonathan Cruz"
date: "2024-02-15"
output: html_document
---

# Project 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(knitr)
```
## Load and Clean
load the text file and remove the data that will not be used for this project

```{r}
str <- read.csv(file="tournamentinfo.txt", header=FALSE, sep="|")
str <- str |> filter(!str_detect(V1, "^-+$"))
str <- str[3 : nrow(str), ]
```

## Isolation
will isolate data based on it's regex matches and filter rows that contains such matches
```{r}
names_collecter <- str |> filter(str_detect(V2, "[A-Z] [A-Z]"))
state_collecter <- str |> filter(str_detect(V1, "[A-Z][A-Z]"))
points_collector <-  str |> filter(str_detect(V3, "\\d+\\.\\d")) 
pre_rating_collector <- str |> filter(str_detect(V2, "\\d+"))
pre_rating_collector <- pre_rating_collector |> mutate(V2 =str_extract(V2,":\\s+\\d+"))
pre_rating_collector <- pre_rating_collector |> mutate(V2 = str_remove_all(V2, " "))
pre_rating_collector <- pre_rating_collector |> mutate(V2 = str_remove_all(V2, ":"))
```

## Table Without Calculations
here will create our table that doesnt not have the avg_pre_rating calculations yet

```{r}
tab <- tibble(player_name = names_collecter$V2, state = state_collecter$V1, total_points = points_collector$V3, pre_rating = pre_rating_collector$V2)
tab <- transform(tab, pre_rating = as.integer(pre_rating))

kable(tab)
```

## Calculate Average Pre Rating
will calculate average_pre_rating

```{r}
avg_pre_vector <- c()
for(i in 1:nrow(points_collector)) {
    count = 0
    adder = 0
    for (j in 5 : ncol(points_collector) - 1){
        if(str_detect(points_collector[i,j] ,"L|D|W") ){
          index_str = str_extract(points_collector[i,j], "\\d+" )
          index = strtoi(index_str)
          adder = adder + tab[index, "pre_rating"]
          count <- count + 1
        }
    }
    avg_pre_vector <- append(avg_pre_vector, round(adder / count))
}

kable(avg_pre_vector)

```

## Attach Calculations and Write CSV
```{r}
tab <- tab |> mutate(avg_pre_rating = avg_pre_vector)

write_csv(tab, file = "results.csv")
kable(tab)
```
