---
title: "Assignment 2"
author: "Jonathan Cruz"
date: "2024-02-02"
output: html_document
---

# MOVIE RATINGS AVERAGE

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Load Libraries
```{r}
library(RMySQL)
library(hrbrthemes)
library(tidyverse)
library(knitr)
```

# Load Table From Database
This extract the values for our rating by connecting to a local database in my machine, I provided an alternate csv method of loading the movie ratings table through csv, with the csv file.
```{r}
#Here i will extract results from database 
mysqlconnection = dbConnect(RMySQL::MySQL(), dbname='movie_ratings', host='localhost', port=3306, user='root', password='Uussd0vtusuZ')
dbListTables(mysqlconnection)
result = dbSendQuery(mysqlconnection, "select * from ratings")
data.frame = fetch(result, n = 10) 
#Below all NA value are displayed
print(data.frame)
```
## Alterative Load as CSV IF No Connection To DB

As csv we extract null values and must convert them to NA in order for it to function just like reading it from a database would
```{r}

data.frame <- read_csv("movie_ratings.csv")
data.frame[data.frame == 'NULL'] <- NA
data.frame <- data.frame |> mutate_at(c("barbie", "past_lives", "the_holdovers", "anatomy_of_fall", "poor_things", "air"), as.numeric)

print(data.frame[2:7])

```

## Calculate Average And Compare 


```{r}
data.frame.reduced <- data.frame |> select(barbie : air)
df <- tribble(~movie_name, ~average,)
kable(df)
# in each column we'd like to remove all NA values in order to calculate average rating based on how many people voted for the movie
for (i in colnames(data.frame.reduced)){
     df[nrow(df) + 1,] = list(i, mean( na.omit(data.frame[[i]])))
}

ggplot(df, aes(x = reorder(movie_name, -average), y = average, fill = movie_name)) +  geom_bar(stat = "identity") + theme_ipsum()

```

